---
title: "Final Project"
author: "Brandon Sucher"
date: "12/1/2021"
output: word_document
---

```{r setup, warning=FALSE, message=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
here::here()
library(Hmisc)
library(flextable)
library(wesanderson)
library(patchwork)
```

# Part 1

This function runs a simulation where the user puts in a sample size, theta (probability of success), and a confidence interval method (Wilson, asymptotic, or exact binomial) and gets 10,000 iterations of that confidence interval from a random binomial sample. We use this function to determine which confidence interval is the most accurate by looking at its coverage probability and also the width of the confidence interval in case the coverage probabilities are similar across methods.
```{r}
simsFunction <- function(sampSize, theta, CImethod) {
  sims <- data.frame()
  nSims <- 10000
  for (i in 1:nSims) {
    # This gets our random sample from the binomial distribution
    samp <- rbinom(sampSize, 1, theta)
    # The binconf function gets a binomial CI based on the method we want
    ciResult <- binconf(sum(samp), sampSize, method = CImethod)
    covProb <- theta >= ciResult[, 2] & theta <= ciResult[, 3]
    sims <- rbind(sims, c(covProb, ciResult[, c(2, 3)]))
  }
  colnames(sims) <- c("CovProb", "LowerBound", "UpperBound")
  return(c(round(mean(sims$CovProb), digits = 4),
           round(mean(sims$UpperBound) - mean(sims$LowerBound), digits = 4)))
}
```

Now, we use our function to test different sample sizes and thetas to find which confidence interval method is the best.
```{r}
sampSizes <- c(20, 40, 100)
thetas <- c(0.05, 0.15)
methods <- c("wilson", "exact", "asymptotic")
simResults <- data.frame()
for (i in 1:length(methods)) {
  for (j in 1:length(thetas)) {
    for (k in 1:length(sampSizes)) {
      simResults <- rbind(simResults,
                          c(methods[i], thetas[j], sampSizes[k],
                            simsFunction(sampSizes[k], thetas[j], methods[i])))
    }
  }
}
colnames(simResults) <- c("Method", "Theta", "Sample_Size", "Cov_Prob",
                          "CI_Width")
# Capitalize is in the Hmisc package and capitalizes the
# first letter of each confidence interval method
simResults$Method <- capitalize(simResults$Method)
```

\pagebreak
## Coverage probability and CI width tables

### Coverage Probability
```{r}
table05 <- simResults[simResults$Theta == 0.05, c(1, 3, 4)]
table15 <- simResults[simResults$Theta == 0.15, c(1, 3, 4)]
# Reshape changes our table from long to be wider and
# make it easier to read when we make it into a table
table05 <- reshape(table05, idvar = "Method", timevar = "Sample_Size",
                   direction = "wide")
table15 <- reshape(table15, idvar = "Method", timevar = "Sample_Size",
                   direction = "wide")
covProb <- rbind(table05, table15)
colnames(covProb) <- c("Method", "20", "40", "100")
Theta <- c(0.05, "", "", 0.15, "", "")
covProb <- cbind(Theta, covProb)

ft <- flextable(covProb, cwidth = 1)
# Header row lets us put the Sample Size header over that
# part of the table while further down, header lines looks
# more like a title
ft <- add_header_row(ft, values = c("", "Sample Size"),
                     colwidths = c(2, 3))
ft <- bold(ft, j = c(1, 2))
ft <- add_header_lines(ft, values =
      "Coverage Probability for Each Sample Size, Theta, and Method")
ft <- theme_box(ft)
ft
```

\pagebreak
### Confidence Interval width
```{r}
table05 <- simResults[simResults$Theta == 0.05, c(1, 3, 5)]
table15 <- simResults[simResults$Theta == 0.15, c(1, 3, 5)]
table05 <- reshape(table05, idvar = "Method", timevar = "Sample_Size",
                   direction = "wide")
table15 <- reshape(table15, idvar = "Method", timevar = "Sample_Size",
                   direction = "wide")
CIwidth <- rbind(table05, table15)
colnames(CIwidth) <- c("Method", "20", "40", "100")
Theta <- c(0.05, "", "", 0.15, "", "")
CIwidth <- cbind(Theta, CIwidth)

ft <- flextable(CIwidth, cwidth = 1.1)
ft <- add_header_row(ft, values = c("", "Sample Size"),
                     colwidths = c(2, 3))
ft <- bold(ft, j = c(1, 2))
ft <- add_header_lines(ft, values =
      "Confidence Interval Width for Each Sample Size, Theta, and Method")
ft <- theme_box(ft)
ft
```

\pagebreak
## Plots

### Coverage probability plot
```{r, fig.width=6}
simResults$Sample_Size <- as.factor(simResults$Sample_Size)
# Need to relevel to get the character values into numeric order
levels(simResults$Sample_Size) <- c("20", "40", "100")
simResults$Cov_Prob <- as.numeric(simResults$Cov_Prob)
simResults$CI_Width <- as.numeric(simResults$CI_Width)

ggplot(simResults, aes(x = Sample_Size, y = Cov_Prob, col = Method)) +
  geom_point(size = 4, aes(shape = Method)) +
  scale_color_manual(values = wes_palette("FantasticFox1")) +
  scale_shape_manual(values = c(1, 0, 4)) +
  facet_wrap(~Theta) +
  labs(x = "Sample Size", y = "Coverage Probability",
       title =
      "Coverage Probability based on sample size, theta, and CI method")
```

\pagebreak

### Confidence interval width plot
```{r, fig.width=6}
ggplot(simResults, aes(x = Sample_Size, y = CI_Width, col = Method)) +
  geom_point(size = 4, aes(shape = Method)) +
  facet_wrap(~Theta) +
  scale_color_manual(values = wes_palette("FantasticFox1")) +
  scale_shape_manual(values = c(1, 0, 4)) +
  labs(x = "Sample Size", y = "Average Width of Confidence Interval",
       title = "Width of CI based on sample size, theta, and CI method")
```

# Part 2

I would like to use the exact binomial interval because it has the greatest coverage probability.

\pagebreak

# Part 3

In this section, we apply our chosen exact binomial interval to a mortality data set with a rare disease. We create the confidence interval for the all individuals and then look into a difference by gender. The table below displays the results.
```{r}
dat <- read.csv("../../OneDrive - Duke University/BIOS_721/mortality_data.csv")
# Take out incorrect values
dat$sex[dat$sex == "-999"] <- NA

# Get CI for all individuals and then by gender
overall <- data.frame(binconf(sum(dat$death),
                              nrow(dat),
                              method = "exact"))
male <- data.frame(binconf(sum(dat$death[dat$sex == "M"], na.rm = TRUE),
                           nrow(dat[dat$sex == "M", ]),
                           method = "exact"))
female <- data.frame(binconf(sum(dat$death[dat$sex == "F"], na.rm = TRUE),
                             nrow(dat[dat$sex == "F", ]),
                             method = "exact"))
allCI <- rbind(overall, male, female)
allCI <- round(allCI, digits = 4)
allCI <- allCI[, c(2, 3)]
Sex <- c("Both", "Male", "Female")
allCI <- cbind(Sex, allCI)
colnames(allCI) <- c("Sex", "Lower Bound", "Upper Bound")

ft <- flextable(allCI, cwidth = 1.25)
ft <- add_header_lines(ft, values = "Proportion of Deaths CI By Sex")
ft <- bold(ft, j = 1)
ft <- theme_box(ft)
ft
```


